---
title: "Calculate frugivores' contribution"
author: "Elena Quintero"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r, message=F}
library(here)
library(tidyverse)
library(magrittr)
library(reshape2)
library(ggrepel)
library(patchwork)

theme_set(theme_minimal())
```

## FRUGIVORES (aggregate data by frugivores)

Load datasets:

```{r}
frug.ints <- read.csv(here("data/frugivores_interactions.csv"))

ind.level.df <- read.csv(here("data/node.level.selection.csv"))

#Net colors:
mycols <- as.character(ind.level.df$cols_continent3)
names(mycols) <- as.character(ind.level.df$plant_sp)

mycols.cont <- as.character(ind.level.df$cols_continent4)
names(mycols.cont) <- as.character(ind.level.df$continent)
```

Accumulated consumption by frugivores in the net:

```{r}
frug.accum.ints <- frug.ints %>% 
  group_by(net_id) %>% 
  mutate(total.ints = sum(ints),
         rank = rank(desc(ints))) %>%
  arrange(net_id, rank) %>%
  mutate(order = order(rank),
         acc_ints = cumsum(ints)) %>% #accumulation in increasing order
  arrange(net_id, desc(order)) %>%
  mutate(inverse_acc_ints = cumsum(ints)) %>% #accumulation in decreasing order
  mutate(core50 = ifelse(inverse_acc_ints > 0.5, "yes", "no"))

glimpse(frug.accum.ints)
```

How many frugivores in the net represent >50% of consumption?

```{r}
d1 <- frug.accum.ints %>%
  group_by(net_id, core50) %>%
  summarise(n=n()) %>%
  pivot_wider(names_from = "core50", values_from = "n") %>%
  mutate(per.core.frug = yes/(yes+no))

summary(d1$per.core.frug)
sd(d1$per.core.frug)
hist(d1$per.core.frug)
```

De media menos del 20% (18.6%) de los frugivores realizan más de la mitad de las interacciones/visitas (rango: min = 2% - 50%)


Calculate bodymass value relative to each network (study site):

```{r}
frug.accum.ints %<>% group_by(net_id) %>%
  mutate(bmass.scaled = scale(bmass))

hist(frug.accum.ints$bmass.scaled)
```


Plot accumulation of interactions:

```{r fig.width=8, fig.height=9}
frug.accum.ints %>%
  filter(!net_id %in% c("08_01", "12_01", "12_06", "12_07")) %>%
  ggplot(aes(x = order, y = ints, group = net_id)) + 
  geom_line(color = "grey70") +
  geom_point(aes(fill = plant_sp, color = core50, size = bmass.scaled), pch = 21) +
  scale_fill_manual(values = mycols) +
  scale_color_manual(values = c("white", "black")) +
  facet_wrap(~fct_reorder(plant_sp, plant_plot_rank), scales = "free", ncol = 4) +
  theme(legend.position = "none", axis.text.x = element_blank(),
        strip.text = element_text(face = "italic"), 
        panel.border = element_rect(linewidth = 0.2, fill = NA),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(x = "Frugivore species ranked by contribution", y = "Contribution to total interactions")
  #caption = "Black contour represent frugivore species that aggregated perform >50% of interactions"

ggsave(here("figs/frugivore_ints_accumulated.pdf"), width = 8, height = 9)
```


Plot for all frugivores in the net the % of interactions it held vs. % of plants it interacted with.

```{r}
frug.accum.ints %>% 
  filter(!net_id %in% c("08_01", "12_01", "12_06", "12_07")) %>%
  ggplot(aes(x = ints, y = per.plants, 
             color = continent,
             label = frug_sp, label2 = net_id)) + 
  geom_point(aes(size = bmass.scaled), alpha = 0.6) +
  scale_color_manual(values = mycols.cont) +
  scale_size_continuous(breaks = c(-1, 0, 1.5, 3, 5)) +
  theme(panel.border = element_rect(fill = NA, linewidth = 0.5)) +
  labs(y = "Proportion of plants in the network visited by each frugivore", 
       x = "Frugivore's contribution to the overall interactions in the network", 
       color = "Region", size = "Body mass \nz-score") 

ggsave(here("figs/interaction_distribution_frugs.pdf"), height = 5, width = 7)
```


Model to test relationship:

```{r}
hist(frug.accum.ints$ints)
hist(frug.accum.ints$per.plants)
hist(log(frug.accum.ints$ints))
hist(log(frug.accum.ints$per.plants))

plot(log(frug.accum.ints$ints), log(frug.accum.ints$per.plants))

data.model <- frug.accum.ints %>% 
  filter(!net_id %in% c("08_01", "12_01", "12_06", "12_07")) %>%
  mutate(net_id = as.factor(net_id)) 

#simple correlations
cor.test(data.model$ints, data.model$per.plants, method = "spearman")
cor.test(data.model$ints, data.model$bmass.scaled, method = "spearman")
cor.test(data.model$per.plants, data.model$bmass.scaled, method = "spearman")
```

