---
title: "Estimate individual resource specialisation"
author: "Francisco Rodriguez-Sánchez & Elena Quintero"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r, message=F}
library(here)
library(tidyverse)
library(magrittr)
library(reshape2)
library(ggrepel)
library(patchwork)

theme_set(theme_bw())
```

Idea based on: *Bolnick, D. I., Svanbäck, R., Fordyce, J. A., Yang, L. H., Davis, J. M., Hulsey, C. D., & Forister, M. L. (2003). The Ecology of Individuals: Incidence and Implications of Individual Specialization. The American Naturalist, 106(1), 1–28.*


Roughgarden (1972) suggested that the total niche width of a population (TNW) can be broken down into two components: the variation in resource use within individuals (within-individual component, WIC), and the variance between individuals (between-individual component, BIC) so that TNW = WIC + BIC (*Bolnick et al. 2002 Ecology*)

Test intra-individual variation in relation to total niche width. See if INTRA-individual variation is higher than INTRA-individual variation as seen in several previous studies.


Load networks:

```{r}
nets_names <- list.files(path = here("networks/standardized/"), pattern = "_int")

nets <- list()

for (i in 1:length(nets_names)){
  net_file <- paste0("networks/standardized/", nets_names[[i]])
  net <- read.csv(here(net_file))
  
  if(ncol(net)<2){
    net <- read.csv(here(net_file), sep = ";") 
  } else
  {}  

  net %<>% column_to_rownames("ind")
  nets[[i]] <- net
  names(nets)[i] <- nets_names[[i]]
}

```


## Shannon index approximation to WIC/TNW.

Original Roughgarden’s WIC/TNW was limited to continuous diet data. But Roughgarden (1979:510) proposes a measure that uses the Shannon-Weaver index as a proxy for variance. 

```{r}
source(here("functions/functions_shannon.R"))
```


Loop for all our nets:

```{r, results = 'hide'}
ind.specialization <- data.frame()

for(i in 1:length(nets)){
  net_id = names(nets[i]) %>% str_sub(3, 7)

  net <- nets[[i]]

  wic_net = wic(net)
  tnw_net = tnw(net)
  ratio = wic(net) / tnw(net)
  names(ratio) <- "WIC/TNW"

  tmp <- cbind(net_id = net_id,
               wic_net,
               tnw_net,
               ratio)

  ind.specialization <- rbind(ind.specialization, tmp)
}

glimpse(ind.specialization)
write_csv(ind.specialization, here("data/WIC_TNW.csv"))

#ind.specialization <- read.csv(here("data/WIC_TNW.csv"))
```


Add network info:

```{r}
net_cols <- read_csv(here("data/net_colors.csv"))

cols.plants <- as.character(net_cols$cols_continent3)
names(cols.plants) <- as.character(net_cols$plant_sp)

cols.cont <- as.character(net_cols$cols_continent4)
names(cols.cont) <- as.character(net_cols$continent)

plot_cols <- net_cols %>% dplyr::select(plant_sp, continent, plant_plot_rank)

ind_nets_ID <- read.csv(here("data/net.level.selection.csv")) %>%
  filter(type == "ind") %>%
  left_join(plot_cols)

ind.specialization %<>% 
  left_join(ind_nets_ID) %>%
  filter(!net_id %in% c("08_01", "12_01", "12_06", "12_07")) # remove small incomplete nets
```

Distribution of WIC/TNW:

```{r}
hist(ind.specialization$WIC.TNW)

summary(ind.specialization$WIC.TNW)
```


boxplot of WIC/TNW Shannon:

```{r plot.height = 3, plot.width = 6}
pos <- position_jitter(height = 0.3, seed = 2)

ggplot(ind.specialization, aes(y = continent, x = WIC.TNW, 
                     color = continent, fill = continent)) +
  geom_boxplot(alpha = .2, outlier.shape = NA, width = 0.4) + 
  geom_jitter(size = 3, position = pos, shape = 21, color = "grey20") + 
  ggrepel::geom_text_repel(aes(label = net_n),
                           size = 2, position = pos) +
  scale_color_manual(values = cols.cont) +
  scale_fill_manual(values = cols.cont) +
  scale_x_continuous(limits = c(0,1)) +
  theme(legend.position = "none",
        panel.border = element_rect(linewidth = 0.5, fill = NA)) +
  labs(x = "WIC/TNW", y  = NULL)  

#ggsave(here("figs/WIC_TNW.pdf"), width = 6, height = 3)
```


Correlate WIC with TNW:

```{r}
ggplot(ind.specialization, aes(x = TNW, y = WIC, 
                     color = continent, fill = continent, 
                     size = number.of.species.HL)) +
  geom_point(shape = 21, color = "grey20") + #size = 3
  ggrepel::geom_text_repel(aes(label = net_n), size = 2)+ 
  scale_color_manual(values = cols.cont) +
  scale_fill_manual(values = cols.cont) +
  theme(#legend.position = "none",
        panel.border = element_rect(linewidth = 0.5, fill = NA)) + 
  geom_abline(slope = 1, linetype = 3) +
  #geom_abline(slope = 0.5, linetype = 3) +
  #scale_x_continuous(limits = c(0,3)) + scale_y_continuous(limits = c(0,2)) +
  scale_y_log10() + scale_x_log10() +
  labs(y = "WIC", x  = "TNW", size = "No. frugivores", color = "Region", fill = "Region") 

#ggsave(here("figs/WIC_vs_TNW_log.pdf"), width = 7, height = 5)
```


Test to determine differences among geographic regions:
```{r}
library(lme4)

hist(ind.specialization$WIC.TNW)

m1<- glmmTMB::glmmTMB(`WIC.TNW` ~ continent + (1|study), data = ind.specialization)
summary(m1)
DHARMa::simulateResiduals(m1, plot =T)

library(modelbased)
estimate_contrasts(m1, contrast = "continent")
```



