---
title: "Comparison species and individual based networks"
author: "Elena Quintero"
date: "`r Sys.Date()`"
output: html_document
---

```{r include = F}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(echo = FALSE)
```


```{r, message=F}
library(here)
library(tidyverse)
library(magrittr)
library(ggfortify)
library(bipartite)
library(igraph)
library(tnet)
library(patchwork)

theme_set(theme_bw())
```

Load observed net-level metrics:

```{r}
net.level <- read.csv(here("data/net.level.selection.csv")) %>%
  mutate(type = recode(type, "ind" = "ind-based", "sp" = "sp-based"))

#glimpse(net.level)
```


Load null model derived net-level metrics:

```{r}
nulls.s <- read_rds(here("data/net_level_nulls_swap.rds")) |> 
  rename(centralization.w = centralization_w)

nulls.p <- read_rds(here("data/net_level_nulls_pat.rds")) |> 
  rename(centralization.w = centralization_w)

nulls.v <- read_rds(here("data/net_level_nulls_vazquez.rds")) |> 
  rename(centralization.w = centralization_w)
```

Select only metrics of interest:

```{r}
metrics.of.interest <- c("connectance", 
                         "weighted.NODF", 
                         "M", 
                         "interaction.evenness",
                         "centralization.w",
                         "assortativity")
```


Convert dataframes to long format:

```{r}
obs.metrics.long <- net.level |> 
  select(net_id, net_n, net_size, type_long = type, all_of(metrics.of.interest)) |> 
  pivot_longer(cols = -c(1:4), names_to = "metric", values_to = "value_obs") |> 
  mutate(type = ifelse(type_long == "ind-based", "ind", "sp")) |> 
  mutate(metric2 = as.factor(recode(metric, 
                         "M" = "Modularity",
                         "assortativity" = "Assortativity",
                         "centralization.w" = "Centralization",
                         "connectance" = "Connectance",
                         "interaction.evenness" = "Interaction evenness",
                         "weighted.NODF" = "Weighted NODF")) |> 
                    fct_relevel("Connectance", "Weighted NODF", "Modularity", "Interaction evenness",
                                "Assortativity", "Centralization"))
#630 observations

null.metrics.long.s <- nulls.s |> 
  ungroup() |> 
  select(net_code, net_n, net_size, type, all_of(metrics.of.interest)) |> 
  pivot_longer(cols = -c(net_code, net_n, net_size, type),
               names_to = "metric", values_to = "value") |> 
  mutate(null = "Swap")

null.metrics.long.p <- nulls.p |> 
  ungroup() |> 
  select(net_code, net_n, net_size, type, all_of(metrics.of.interest)) |> 
  pivot_longer(cols = -c(net_code, net_n, net_size, type),
               names_to = "metric", values_to = "value") |> 
  mutate(null = "Patefield")

null.metrics.long <- full_join(null.metrics.long.s, null.metrics.long.p)
#1,260,000 observations

null.metrics.long.selec <- null.metrics.long |> 
  mutate(null_metric = paste(metric, null)) |> 
  filter(null_metric %in% c("connectance Patefield", "M Patefield", "assortativity Patefield",
                            "centralization.w Patefield", "interaction.evenness Patefield", "weighted.NODF Patefield")) |> 
  mutate(metric2 = as.factor(recode(metric, 
                         "M" = "Modularity",
                         "assortativity" = "Assortativity",
                         "centralization.w" = "Centralization",
                         "connectance" = "Connectance",
                         "interaction.evenness" = "Interaction evenness",
                         "weighted.NODF" = "Weighted NODF")) |> 
                    fct_relevel("Connectance", "Weighted NODF", "Modularity", "Interaction evenness",
                                "Assortativity", "Centralization"))
```

Check null-models metrics distribution:

```{r fig.width=10, fig.height=8, include=FALSE}
plot.fun <- function(metric, null_type) {
  null.metrics.long |>
    filter(null == {{null_type}}) |> 
    filter(metric == {{metric}}) |>
  ggplot(aes(x = value, fill = type)) +
  #geom_density(alpha = 0.5) +
  geom_histogram() +
  geom_abline() +
  scale_fill_manual(values = c("grey20", "grey80")) +
  theme(legend.position = "none",
        axis.text.y = element_blank()) +
  labs(y = NULL, x = NULL, title =  {{metric}}) + 
  facet_wrap(~net_code, scales = "free") +
  geom_vline(aes(xintercept = get(metric)), color = "goldenrod", lwd=1.2, 
             data = net.level)
  }

plot.fun("connectance", "Patefield")
plot.fun("weighted.NODF", "Swap")
plot.fun("interaction.evenness", "Swap")
plot.fun("M", "Swap")
plot.fun("assortativity", "Swap")
plot.fun("centralization.w", "Swap")
```


Summarize null model metrics:

```{r}
null.data.sum.s <- nulls.s |> 
  select(type, net_id, iter, all_of(metrics.of.interest)) |> 
  pivot_longer(cols = -c(1:3), names_to = "metric", values_to = "value") |> 
  group_by(type, net_id, metric) |> 
  summarise(mean_nulls = mean(value),
            sd_nulls = sd(value)) |> 
  mutate(null = "Swap")

null.data.sum.p <- nulls.p |> 
  select(type, net_id, iter, all_of(metrics.of.interest)) |> 
  pivot_longer(cols = -c(1:3), names_to = "metric", values_to = "value") |> 
  group_by(type, net_id, metric) |> 
  summarise(mean_nulls = mean(value),
            sd_nulls = sd(value)) |> 
  mutate(null = "Patefield")

null.data.sum <- full_join(null.data.sum.s, null.data.sum.p)|> 
  left_join(obs.metrics.long) |> 
  mutate(z_score = (value_obs - mean_nulls) / sd_nulls,
         diff_nulls = (value_obs - mean_nulls))

null.data.sum.selec <- null.data.sum |> 
  mutate(null_metric = paste(metric, null)) |> 
  filter(null_metric %in% c("connectance Patefield", "M Patefield", "assortativity Patefield",
                            "centralization.w Patefield", "interaction.evenness Patefield", "weighted.NODF Patefield")) |> 
  mutate(metric = as.factor(recode(metric, 
                         "M" = "Modularity",
                         "assortativity" = "Assortativity",
                         "centralization.w" = "Centralization",
                         "connectance" = "Connectance",
                         "interaction.evenness" = "Interaction evenness",
                         "weighted.NODF" = "Weighted NODF")) |> 
                    fct_relevel("Connectance", "Weighted NODF", "Modularity", "Interaction evenness",
                                "Assortativity", "Centralization"))
```


Plot to compare network types:

```{r fig.width = 10, fig.height = 5}
p1 <- null.data.sum.selec |> 
  ggplot(aes(x = value_obs, fill = type_long, color = type_long)) + 
  geom_density(alpha = 0.4) + 
  facet_wrap(~metric, scales = "free", nrow = 1) + 
  scale_fill_manual(values = c("goldenrod", "grey40")) +
  scale_color_manual(values = c("goldenrod", "grey40")) +
  #scale_fill_manual(values = c("grey40", "#2F8AC4")) +
  theme(panel.background = element_rect(colour = "black"),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank()) + 
  labs(y = NULL, x = NULL, title = "A. Observed values",
       color = "Network\ntype", fill = "Network\ntype")

p2 <- null.data.sum.selec |> 
  ggplot(aes(x = diff_nulls, fill = type, color = type)) + 
  geom_density(alpha = 0.4) + 
  facet_wrap(~metric, scales = "free", nrow = 1) + 
  scale_fill_manual(values = c("goldenrod", "grey40")) +
  scale_color_manual(values = c("goldenrod", "grey40")) +
  #scale_fill_manual(values = c("grey40", "#2F8AC4")) +
  theme(legend.position = "none",
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(),
        # panel.background = element_rect(colour = "black"), 
        # strip.background = element_rect(colour = "white"),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank()) + 
  labs(y = NULL, x = NULL, title = "B. Deviation from null models")

p1 / p2 + plot_layout(guides = "collect") + 
  plot_annotation(caption = 'Deviation from null model is in respect to Swap algorithm,\n except connectance which is in respect to Patefield algorithm.')

#ggsave(here("figs/Fig_2_net_level_comparison.pdf"), width = 10, height = 5)

```


Summary metrics:

```{r}
sum.tb <- null.data.sum.selec |> 
  group_by(metric, type) |> 
  summarise(obs_mean = round(mean(value_obs), 2),
            obs_sd = round(sd(value_obs), 2),
            nulls_mean = round(mean(mean_nulls), 2),
            nulls_sd = round(mean(sd_nulls), 2),
            nulls_diff = round(mean(diff_nulls), 2),
            nulls_diff_sd = round(sd(diff_nulls), 2),
            nulls_zs = round(mean(z_score), 2)) |> 
  mutate(obs = paste0(obs_mean, " ± ", obs_sd),
         nulls = paste0(nulls_mean, " ± ", nulls_sd),
         diff = paste0(nulls_diff, " ± ", nulls_diff_sd)) |> 
  select(metric, type, obs, nulls, diff)

sum.tb

#write_csv(sum.tb, here("tables/net_metrics_summary.csv"))
```


Plot comparing metrics by network one by one:

```{r, fig.height=6, fig.width=9} 
plot.null <- function(metric) {
  null.metrics.long.selec |>
    mutate(type = recode(type, "ind" = "individual-based networks", "sp" = "species-based networks")) |> 
    # filter(null == {{null_type}}) |> 
    filter(metric2 == {{metric}}) |>
    ggplot(aes(x = as.factor(net_n), y = value)) +
    ggdist::stat_pointinterval(.width = 0.95, point_size = 0.8, linewidth = 0.5, color = "grey30") + 
    #geom_boxplot(outliers = FALSE) +
    geom_point(aes(x = as.factor(net_n), y = value_obs), color = "goldenrod", pch =8, size = 2,
               data = obs.metrics.long |> filter(metric2 == {{metric}}) |>
                 mutate(type = recode(type, "ind" = "individual-based networks", 
                                      "sp" = "species-based networks"))) +
    theme_minimal() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0.75), axis.title = element_blank(), 
          panel.background = element_rect(colour = "grey50", fill = NA)) + 
    facet_wrap(~type, scales = "free", ncol = 1) 
    #labs(title = paste0({{metric}}, " (",{{null_type}}, ")"), x = "net code", y = NULL)
}

ps1 <- plot.null("Connectance") + labs(title = "A. Connectance")
ps2 <- plot.null("Weighted NODF") + labs(title = "B. Weigthed NODF")
ps3 <- plot.null("Modularity") + labs(title = "C. Modularity")
ps4 <- plot.null("Interaction evenness") + labs(title = "D. Interaction evenness")
ps5 <- plot.null("Assortativity") + labs(title = "E. Assortativity")
ps6 <- plot.null("Centralization") + labs(title = "F. Centralization")

ps1 / ps2 
ps3 / ps4
ps5 / ps6

ggsave(filename = here("figs/Fig_SM_A.png"), plot = ps1, width = 7, height = 5)
ggsave(filename = here("figs/Fig_SM_B.png"), plot = ps2, width = 7, height = 5)
ggsave(filename = here("figs/Fig_SM_C.png"), plot = ps3, width = 7, height = 5)
ggsave(filename = here("figs/Fig_SM_D.png"), plot = ps4, width = 7, height = 5)
ggsave(filename = here("figs/Fig_SM_E.png"), plot = ps5, width = 7, height = 5)
ggsave(filename = here("figs/Fig_SM_F.png"), plot = ps6, width = 7, height = 5)
```

Alternative fig 2:

```{r fig.height=6, fig.width=9}
p3 <- obs.metrics.long |>
  rename(value = value_obs) |> 
  mutate(basis_of_value = "obs") |> 
  full_join(null.metrics.long.selec)  |> 
  mutate(basis_of_value = if_else(is.na(basis_of_value), "null", basis_of_value)) |> 
  mutate(basis_of_value = fct_relevel(basis_of_value, c("obs", "null")))  |> 
  ggplot(aes(x = type, y = value, color = basis_of_value, fill = basis_of_value)) +
  geom_boxplot(outliers = FALSE, alpha = 0.4) +
  scale_color_manual(values = c("goldenrod", "grey40")) + 
  scale_fill_manual(values = c("goldenrod", "grey40")) +
  facet_wrap(~metric2, scales = "free", nrow = 1) +
  labs(x = NULL, y = NULL)

p4 <- null.metrics.long.selec |> 
  mutate(basis_of_value = "difference\n(obs - null)") |> 
  left_join(obs.metrics.long) |> 
  mutate(diff = (value_obs - value)) |> 
  ggplot(aes(x = type, y = diff, color = basis_of_value, fill = basis_of_value)) +
  geom_boxplot(outliers = FALSE, alpha = 0.4) +
  # scale_color_manual(values = "#009999") + 
  # scale_fill_manual(values = "#009999") +
  facet_wrap(~metric2, scales = "free", nrow = 1) +
  labs(x = NULL, y = NULL)

p3 / p4 + plot_layout(guides = "collect")
```

2nd Alternative fig 2:

```{r fig.height=6, fig.width=9}
p3 <- obs.metrics.long |>
  rename(value = value_obs) |> 
  mutate(basis_of_value = "obs") |> 
  mutate(type = recode(type, "ind" = "Ind", "sp" = "Sp")) |> 
  ggplot(aes(x = type, y = value)) +
  ggdist::stat_pointinterval(aes(color = type, fill = type), alpha = 0.6, .width = c(0.66, 0.95)) +
  #geom_boxplot(outliers = FALSE, alpha = 0.4) + # color = "goldenrod", fill =  "goldenrod"
  scale_color_manual(values = c("goldenrod", "grey40")) + 
  scale_fill_manual(values = c("goldenrod", "grey40")) +
  facet_wrap(~metric2, scales = "free", nrow = 1) +
  labs(x = NULL, y = NULL, subtitle = "A. Observed values") + 
  theme(legend.position = "none", strip.text = element_text(size = 7))

# model_labels <- data.frame(metric2 = unique(null.metrics.long.selec$metric2),
#          label = c("", "*", "*", "", "*", "*")) |> 
#   mutate(metric2 = metric2 |> fct_relevel("Connectance", "Weighted NODF", "Modularity",
#                                           "Interaction evenness",
#                                           "Assortativity", "Centralization"))

p4 <- null.metrics.long.selec |> 
  mutate(basis_of_value = "difference\n(obs - null)") |> 
  left_join(obs.metrics.long) |> 
  mutate(diff = (value_obs - value)) |> 
  mutate(type = recode(type, "ind" = "Ind", "sp" = "Sp")) |> 
  ggplot(aes(x = type, y = diff)) +
  ggdist::stat_pointinterval(aes(color = type, fill = type), alpha = 0.6, .width = c(0.66, 0.95)) +
  #geom_boxplot(aes(color = type, fill = type), outliers = FALSE, alpha = 0.4) +# color = "grey40", fill =  "grey40"
  scale_color_manual(values = c("goldenrod", "grey40")) + 
  scale_fill_manual(values = c("goldenrod", "grey40")) +
  facet_wrap(~metric2, scales = "free", nrow = 1) +
  labs(x = NULL, y = NULL, subtitle = "B. Deviation from random expectation")+ 
  theme(legend.position = "none", strip.text = element_text(size = 7)) 
  # geom_text(x = 1.5, y = c(0, 0.28, 0.01, 0, 0.21, -0.03), aes(label = label), data = model_labels)

p3 / p4 + plot_layout(guides = "collect")

#ggsave(here("figs/Fig_2_net_level_comparison.png"), width = 7, height = 4)
```

3rd alternative to fig:

```{r fig.height=6, fig.width=9}
p5 <- obs.metrics.long |> 
  mutate(basis_of_value = "observed") |> 
  ggplot(aes(y = type, x = value_obs, color = type, fill = type)) +
  geom_boxplot(outliers = FALSE, alpha = 0.4) +
  scale_fill_manual(values = c("goldenrod", "grey40")) +
  scale_color_manual(values = c("goldenrod", "grey40")) +
  facet_wrap(~metric2, scales = "free", nrow = 1) +
  labs(x = NULL, y = NULL) + 
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank())

p6 <- null.metrics.long.selec |> 
  mutate(basis_of_value = "difference\n(obs - null)") |> 
  left_join(obs.metrics.long) |> 
  mutate(diff = (value_obs - value)) |> 
  ggplot(aes(y = type, x = diff, color = type, fill = type)) +
  geom_boxplot(outliers = FALSE, alpha = 0.4) +
  scale_fill_manual(values = c("goldenrod", "grey40")) +
  scale_color_manual(values = c("goldenrod", "grey40")) +
  facet_wrap(~metric2, scales = "free", nrow = 1) +
  labs(x = NULL, y = NULL) + 
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank())

(p1 / p5 / p2 / p6) + plot_layout(heights = c(4,1,4,1), guides = "collect")
```

Model comparisons:

```{r}
library(glmmTMB)
library(DHARMa)

tmp <- null.metrics.long.selec |> 
  mutate(basis_of_value = "difference\n(obs - null)") |> 
  left_join(obs.metrics.long) |> 
  mutate(diff = (value_obs - value))

m1 <- glmmTMB::glmmTMB(diff ~ type + (1|net_code), data = tmp |> filter(metric == "connectance"))
summary(m1)
simulateResiduals(m1, plot = T)

m2 <- glmmTMB::glmmTMB(diff ~ type + (1|net_code), data = tmp |> filter(metric == "weighted.NODF"))
summary(m2)
simulateResiduals(m2, plot = T)

m3 <- glmmTMB::glmmTMB(diff ~ type + (1|net_code), data = tmp |> filter(metric == "M"))
summary(m3)
simulateResiduals(m3, plot = T)

m4 <- glmmTMB::glmmTMB(diff ~ type + (1|net_code), data = tmp |> filter(metric == "interaction.evenness"))
summary(m4)
simulateResiduals(m4, plot = T)

m5 <- glmmTMB::glmmTMB(diff ~ type + (1|net_code), data = tmp |> filter(metric == "assortativity"))
summary(m5)
simulateResiduals(m5, plot = T)

m6 <- glmmTMB::glmmTMB(diff ~ type + (1|net_code), data = tmp |> filter(metric == "centralization.w"))
summary(m6)
simulateResiduals(m6, plot = T)

rbind(summary(m1)$coefficients$cond,
      summary(m2)$coefficients$cond,
      summary(m3)$coefficients$cond,
      summary(m4)$coefficients$cond,
      summary(m5)$coefficients$cond,
      summary(m6)$coefficients$cond)
```




```{r}
obs.metrics.long2 <- net.level |> 
  select(net_id, net_n, net_size, type, 
         number.of.species.HL, number.of.species.LL,
         connectance, weighted.NODF, M, interaction.evenness,
         centralization.w, assortativity, ) |> 
  pivot_longer(cols = -c(1:6), names_to = "metric", values_to = "value_obs") |> 
  mutate(metric2 = as.factor(recode(metric, 
                         "M" = "Modularity",
                         "assortativity" = "Assortativity",
                         "centralization.w" = "Centralization",
                         "connectance" = "Connectance",
                         "interaction.evenness" = "Interaction evenness",
                         "weighted.NODF" = "Weighted NODF")) |> 
                    fct_relevel("Connectance", "Weighted NODF", "Modularity", "Interaction evenness",
                                "Assortativity", "Centralization"))

ggplot(obs.metrics.long2, aes(x = number.of.species.HL + number.of.species.LL,
                      y = value_obs,
                      color = type)) + 
  geom_point(size = 2.5) + 
  facet_wrap(~metric2, scales = "free_y") +
  scale_color_manual(values = c("goldenrod2", "grey30")) + 
  theme_minimal() + 
  theme(panel.background = element_rect(colour = "grey50", fill = NA)) +
  labs(x = "Number of network nodes", color = NULL, y = NULL)

ggsave(here("Figs/SM_corr.png"), width = 9, height = 5)
```


