---
title: "Networks standarization with posteriors"
author: "Elena Quintero"
date: "`r Sys.Date()`"
output: github_document
---

```{r, message=F}
library(here)
library(tidyverse)
library(network.tools)
library(patchwork)
```

```{r setup, include=FALSE}
knitr::knit_hooks$set(timeit = local({
  now = NULL
  function (before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      # calculate the time difference after a chunk
      res <- difftime(Sys.time(), now, units = "mins")
      # return a character string to show the time
      paste("Time for this code chunk to run:", round(res,
        2), "minutes")
    }
  }})
)
```

# Data

## 1. Specify parameters:

(1) network id, (2) sampling effort and (3) whether controlling for differential sampling effort in plant individuals is needed:

```{r}
params <- read_csv2(here("networks/nets_se/nets_parameters.csv")) |> 
  filter(str_detect(net_code, "ind"))

# List of individual-based networks
net_id <- params$net_id

# List of sampling effort for each individual-based networks
se_net <- params$sampling_effort

# List of whether individuals require controlling by sampling effort
control_needed <- params$necessary_stand

```

## 2. Read network posterior counts:

Load data resulting from Bayes model.

```{r timeit=TRUE, message=FALSE, results='hide'}
files <- list.files(here("networks/nets_post"))
nets <- str_sub(files, 1, 5)
#remove small nets
nets_selec <- nets[! nets %in% c("08_01", "12_01", "12_06", "12_07")]

nets <- list()

for (i in 1:length(net_id)) {
  
  if (net_id[i] %in% nets_selec) {
    
    print(net_id[i])
    
    net.post <- readRDS(here::here(paste0("networks/nets_post/",
                                        net_id[i],
                                        "_post_counts.rds"))) |>
      group_by(Plant, Animal) |>
      mutate(iter = c(1:1000)) |>
      mutate(net_id = net_id[i])

   nets[[i]] <- net.post
   names(nets)[i] <- net_id[i]
  
  }
  else {
  }
}
```


## 3. Read info on sampling effort for each network:

```{r timeit=TRUE, message=FALSE, results='hide'}
nets.se <- list()

for (i in 1:length(net_id)) {
  
  if (control_needed[i] == "YES") {
    
    print(net_id[i])
    
    effort <- read_csv(here::here(paste0("networks/nets_se/",
                                         net_id[i],
                                         "_attr.csv"))) |>
      select(ind, SE = se_net[i]) |> 
      mutate(Plant = as.character(ind))

    nets.se[[i]] <- effort
    names(nets.se)[i] <- net_id[i]
  
  }
  else {
  }
}
```


## 4 Divide posterior counts by sampling effort and round interactions:

Only if necessary controlling by sampling effort. That is, networks where sampling effort was already accounted for or where plant individuals where sampled with the same effort do not require this step.

Then round interactions so that interactions under 0.5 counts are not consiredes to have happened.

```{r timeit=TRUE, message=FALSE, results='hide'}
nets.post.se <- list()

for (i in 1:length(nets)) {
  
    if (control_needed[i] == "YES"){
      
      net.post.se <- nets[[i]] |> 
        left_join(nets.se[[i]], by = "Plant") |> 
        mutate(count_r = round(count)) |> #round posterior
        mutate(count_se = (count_r / SE))
      
      nets.post.se[[i]] <- net.post.se
      names(nets.post.se)[i] <- net_id[i]
      
    }
  
    else {
      
      net.post.se <- nets[[i]] |> 
        mutate(count_r = round(count)) |> #round posterior
        mutate(count_se = count_r)
      
      nets.post.se[[i]] <- net.post.se
      names(nets.post.se)[i] <- net_id[i]
      
    }
}
```

Multiply counts by 10 in network "02_01":

```{r}
nets.post.se[[3]] <- nets.post.se[[3]] |> 
  mutate(count_r = count_r*10,
         count_se = count_se*10)
```



## 5. Standarize matrix

Refer matrix to interaction proportion by dividing all unique interaction by the total no. of interactions (a.k.a. Grand Total Standardization, GTS) - this makes them comparable.

```{r, timeit=TRUE, message=FALSE, results='hide'}

gts <- function(net) {
  
  # calculate total counts per posterior sample
  total.count.iter <- net |> 
    group_by(.draw) |> 
    summarise(total = sum(count_se))
  
  # divide counts by total
  net.std <- net |> 
    left_join(total.count.iter, by = ".draw") |> 
    mutate(count.std = count_se/total)
  
}

# apply gts function to each network at once
nets.std <- lapply(nets.post.se, gts)
names(nets.std) <- names(nets.post.se)

```


## 6. Save matrices

```{r timeit=TRUE, message=FALSE, results='hide'}

for (i in 1:length(nets.std)) {
  if(!is.null(nets.post.se[[i]])){
    
    to_save <- nets.std[[i]] |> 
      select(iter = .draw, 
             Plant, 
             Animal, 
             Visits = count.std)
    
    net_id <- names(nets.std)[i]
    saveRDS(to_save, here(paste0("networks/nets_post_std/", net_id, "_int_std.rds")))
    
    }
}
```


```{r}
##Checking each network iteration sums 1
#a <- readRDS(here(paste0("data/nets_post_std/", net_id, "_int_std.rds")))
# 
# a |> 
#    group_by(iter) |> 
#    summarise(sum(Visits))
```



