---
title: "Networks standarization"
author: "Elena Quintero"
date: "`r Sys.Date()`"
output: github_document
---

```{r, message=F}
library(here)
library(tidyverse)
library(network.tools)
library(patchwork)
```

# Data

## 1. Specify parameters:

(1) network id, (2) sampling effort and (3) whether controlling for differential sampling effort in plant individuals is needed:

```{r}
params <- read_csv2(here("networks/nets_se/nets_parameters.csv"))

# List of networks
net_id <- params$net_id

# List of sampling effort for each individual-based networks
se_net <- params$sampling_effort

# List of whether individuals require controlling by sampling effort
control_needed <- params$necessary_stand

```

## 2. Read individual-based networks:

```{r message=FALSE, results='hide'}
ind.files <- list.files(here("networks/nets_raw"))

nets.ind <- str_sub(ind.files, 1, 5)

#remove small nets
nets.selec.ind <- nets.ind[! nets.ind %in% c("08_01", "12_01", "12_06", "12_07")]

nets.ind <- list()

for (i in 1:length(nets.selec.ind)) {
  
  print(nets.selec.ind[i])
    
    net <- read_csv(here::here(paste0("networks/nets_raw/",
                                        nets.selec.ind[i],
                                        "_int.csv")))

    nets.ind[[i]] <- net
    names(nets.ind)[i] <- nets.selec.ind[i]

}
```


## 2. Read species-based networks:

```{r message=FALSE, results='hide'}
sp.files <- list.files(here("networks/nets_sp_raw"))

nets.sp <- str_sub(sp.files, 1, 8)

#remove small nets
nets.selec.sp <- nets.sp[! nets.sp %in% c("sp_03_01", "sp_03_02",
                                          "sp_03_03", "sp_03_04",
                                          "sp_05_01", "sp_09_01",
                                          "sp_29_12", "sp_31_01",
                                          "sp_44_01")]

nets.sp <- list()

for (i in 1:length(nets.selec.sp)) {
    
    print(nets.selec.sp[i])
    
    net <- read.csv(here::here(paste0("networks/nets_sp_raw/",
                                        nets.selec.sp[i],
                                        ".csv")), sep = ";")

   nets.sp[[i]] <- net
   names(nets.sp)[i] <- nets.selec.sp[i]
  
}
```


## 3. Read info on sampling effort for individual-based network:

```{r message=FALSE, results='hide'}
nets.ind.se <- list()

for (i in 1:length(nets.selec.ind)) {
  
  if (control_needed[i] == "YES") {
    
    print(net_id[i])
    
    effort <- read_csv(here::here(paste0("networks/nets_se/",
                                         nets.selec.ind[i],
                                         "_attr.csv"))) |>
      select(ind, SE = se_net[i]) |> 
      mutate(ind = as.character(ind))

    nets.ind.se[[i]] <- effort
    names(nets.ind.se)[i] <- nets.selec.ind[i]
  
  }
  else {
  }
}
```


## 4 Divide interactions by sampling effort:

Only if necessary controlling by sampling effort. That is, networks where sampling effort was already accounted for or where plant individuals where sampled with the same effort do not require this step.

```{r message=FALSE, results='hide'}
for (i in 1:length(nets.ind)) {
  
    if (control_needed[i] == "YES"){
      
      net.se <- nets.ind[[i]] |> 
        mutate(ind = as.character(ind)) |> 
        pivot_longer(cols = -1, names_to = "animal", values_to = "int") |>
        left_join(nets.ind.se[[i]], by = "ind") |> 
        mutate(int = int / SE) |> 
        pivot_wider(names_from = animal, values_from = int) |> 
        select(-SE)
      
      nets.ind[[i]] <- net.se
      names(nets.ind)[i] <- nets.selec.ind[i]
      
    }
  
    else {

    }
}
```


## 5. Standarize matrix

Refer matrix to interaction proportion by dividing all unique interaction by the total no. of interactions (a.k.a. Grand Total Standardization, GTS) - this makes them comparable.

For individual-based nets:

```{r, message=FALSE, results='hide'}


gts <- function(net) {
  
  net_mat <- net |> column_to_rownames("ind") |>  as.matrix()
  mat_gts <- net_mat/sum(net_mat)
  net_gts <- as.data.frame(mat_gts) %>% rownames_to_column("ind")
  
}

nets.std.ind <- lapply(nets.ind, gts)
names(nets.std.ind) <- names(nets.ind)
```

For species-based nets:

```{r, message=FALSE, results='hide'}
gts.sp <- function(net) {
  
  net_mat <- net |> column_to_rownames("sp") |>  as.matrix()
  mat_gts <- net_mat/sum(net_mat)
  net_gts <- as.data.frame(mat_gts) %>% rownames_to_column("sp")
  
}

nets.std.sp <- lapply(nets.sp, gts.sp)
names(nets.std.sp) <- names(nets.sp)

```


## 6. Save matrices

```{r message=FALSE, results='hide'}
for (i in 1:length(nets.std.ind)) {
    
    write_csv(nets.std.ind[[i]], 
            here(paste0("networks/nets_std/", 
                        names(nets.std.ind)[i], 
                        "_int_std.csv")))
}
```


```{r message=FALSE, results='hide'}
for (i in 1:length(nets.std.sp)) {
    
    write_csv(nets.std.sp[[i]], 
            here(paste0("networks/nets_std/", 
                        names(nets.std.sp)[i], 
                        "_int_std.csv")))
}

```
