---
title: "Sampling completeness"
author: "Elena Quintero"
date: "`r Sys.Date()`"
output: github_document
---

```{r include = F}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```


```{r, message=F}
library(here)
library(tidyverse)
library(magrittr)
library(reshape2)
library(patchwork)
library(ggalt)
library(iNEXT)
library(iNEXT.link)

theme_set(theme_minimal())
```

Load colors and other attributes:
```{r}
net_cols <- read_csv(here("data/net_colors.csv"))

cols.plants <- as.character(net_cols$cols_continent3)
names(cols.plants) <- as.character(net_cols$plant_sp)

cols.cont <- as.character(net_cols$cols_continent4)
names(cols.cont) <- as.character(net_cols$continent)

nets_info <- read_csv(here("data/net.level.selection.csv")) %>%
  left_join(net_cols) %>%
  filter(type == "ind") %>%
  select(-c(connectance:centralization.w, cols_country:cols_continent4)) %>%
  rename(net_id = code_ID)

glimpse(nets_info)
```


Read all networks:

```{r}
#Read list of names
nets_names <- list.files(path = here("networks/"), pattern = "_int")

#Create empty list
nets <- list()

#Add all read files to the list
for (i in 1:length(nets_names)){
  net_file <- paste0("networks/", nets_names[[i]])
  net <- read.csv(here(net_file))
  
  if(ncol(net)<2){
    net <- read.csv(here(net_file), sep = ";") #For those datasets saved with ; separation
  } else
  {}  

  net %<>% mutate(ind = as.factor(ind))#column_to_rownames("ind")
  nets[[i]] <- net #add to list
  names(nets)[i] <- nets_names[[i]] #name the list

  # assign(nets_names[[i]], net)
}
```

Read nets characteristics (name, site etc...)

```{r}
nets_ID <- readxl::read_xlsx(here("data/selected_individual_networks.xlsx")) %>%
  select(net_id = code_ID,
         ref,
         plant_sp = `focus species`,
         pop = pop_site) %>%
  mutate(net_name = paste0(plant_sp, "_", pop)) %>%
  group_by(net_id) %>%
  mutate(net_n = cur_group_id())
```

```{r}
nets.melted <- melt(nets) %>%
  rename(animal = variable) %>%
  rename(int = value) %>%
  mutate(net_id = str_sub(L1, 1,5)) %>%
  left_join(nets_ID, by = "net_id") %>%
  mutate(study_n = as.numeric(str_sub(L1, 1,2)),
         subnet_n = as.numeric(str_sub(L1, 4,5))) %>%
  filter(!net_id %in% c("08_01", "12_01", "12_06", "12_07")) #remove nets with few observations

glimpse(nets.melted)
```

## Calculate sampling completeness

### Using inext

Incidence-frequency data: input data for each assemblage consist of the number of sampling units (T) followed by the observed incidence frequencies (Y1 , Y2 , . .., YS ). When there are N assemblages, input data consist of an S + 1 by N matrix or N lists of species incidence frequencies. The first entry of each column/list must be the total number of sampling units, followed by the species incidence frequencies - this is qualitative, is not considering the abundance of frugivore species, just the number of frugivores detected in each plant.

```{r}
inext_sc  <- data.frame()

for (i in 1:length(unique(nets.melted$study_n))){

  net <- nets.melted %>% filter(study_n == i) 
  nets_se <- list()
  
  for (i in unique(net$subnet_n)){
    subnet <- net %>% filter(subnet_n == i)
    
    if (isTRUE(length(unique(subnet$ind)) >= 3)) {
       
    sp_detected <- as.data.frame(subnet) %>% group_by(animal) %>% 
                       filter(int != 0) %>% count()
    
    nets_se[[i]] <-  c(length(unique(subnet$ind)), 
                        sp_detected$n)
    
    names(nets_se)[i] <- unique(subnet$net_id)
    
    
    }
    else
    {}
                              
  }
  
  nets_se[sapply(nets_se, is.null)] <- NULL

  out <- iNEXT(nets_se, q = 0, datatype = "incidence_freq")

  inext_sc <- rbind(inext_sc, out$iNextEst$size_based %>% filter(Method == "Observed"))
  
}

```

Plot sampling coverage:

```{r}
glimpse(inext_sc)

sc <- inext_sc %>% 
  rename(net_id = Assemblage) %>%
  left_join(nets_info)

glimpse(sc)
  
write_csv(sc, here("data/sampling_coverage_inext.csv"))
```


```{r}
ggplot(inext.sc, aes(x = SC, fill = plant_sp)) + 
  geom_histogram() + 
  scale_fill_manual(values = cols.plants)
```



Species Accumulation Curves for each study:

```{r sac, fig.width = 7}

plot_list = list()

for (i in 1:length(unique(nets.melted$study_n))){

  net <- nets.melted %>% filter(study_n == i) 
  nets_se <- list()
  
  for (z in unique(net$subnet_n)){
    subnet <- net %>% filter(subnet_n == z)
    
    if (isTRUE(length(unique(subnet$ind)) >= 3)) {
       
    sp_detected <- as.data.frame(subnet) %>% group_by(animal) %>% 
                       filter(int != 0) %>% count()
    
    nets_se[[z]] <-  c(length(unique(subnet$ind)), 
                        sp_detected$n)
    
    names(nets_se)[z] <- ifelse(unique(subnet$pop) == "NA",
                                unique(subnet$plant_sp), 
                                unique(subnet$pop))
    }
    else
    {}
                              
  }
  
  nets_se[sapply(nets_se, is.null)] <- NULL

  out <- iNEXT(nets_se, q = 0, datatype = "incidence_freq")
  
  #Plot species diversity
  plot <- ggiNEXT(out, type = 1,
                  color.var = "Assemblage") + 
    theme_bw() +
    labs(title = unique(net$ref),
         x = "Number of individual plants",
         y = "Frugivore species diversity") +
    scale_shape_manual(values = rep(16, length(nets_se))) +
    theme(legend.position = "bottom")

  plot_list[[i]] = plot

  }
```


Save all plots to a pdf (several plots per page)
```{r}
ggsave(
   filename = "figs/sampling_coverage.pdf", 
   plot = gridExtra::marrangeGrob(plot_list, nrow=5, ncol=2),
   width = 10, height = 15
)
```


