---
title: "Individual specialization (WIC/TNW) using network posteriors"
author: "Elena Quintero"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::knit_hooks$set(timeit = local({
  now = NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      # calculate the time difference after a chunk
      res <- difftime(Sys.time(), now, units = "mins")
      # return a character string to show the time
      paste("Time for this code chunk to run:", round(res,
        2), "minutes")
    }
  }})
)
```

```{r, message=F}
library(here)
library(tidyverse)
library(network.tools)
library(ggdist)
library(ggridges)

theme_set(theme_bw())
```


## Data

Read network posteriors:

```{r}
files <- list.files(here("networks/nets_post_std"))

nets <- list()

for (i in 1:length(files)) {
  net.post <- readRDS(here::here(paste0("networks/nets_post_std/", files[i]))) |> 
    mutate(net_id = str_sub(files[i], 1, 5))

  nets[[i]] <- net.post
  names(nets)[i] <- str_sub(files[i], 1, 5)
}

```

## Calculate WIC/TNW metric

Using Shannon approximation

```{r eval=FALSE, timeit = TRUE}
mytime <- Sys.time()

ind.spec <- data.frame()

for (i in 1:length(nets)) {
  
  print(names(nets[i]))
  
  for (j in 1:1000) {
    
    net <- nets[[i]] |> 
      filter(iter == j) 
   
    mat <- long2wide(net) 
    
    ind.sp <- indiv_spec(mat)
    
    tmp <- data.frame(net_id = names(nets)[i],
                      iter = j,
                      WIC = ind.sp$WIC,
                      TNW = ind.sp$TNW,
                      WIC.TNW = ind.sp$IndSpec)
  
    ind.spec <- rbind(ind.spec, tmp)
  }
}

Sys.time()-mytime
```

```{r eval=FALSE}
saveRDS(ind.spec, here("metrics/WIC_TNW_posts.rds"))
```


Get number of plants and frugivores per iteration:

```{r}
nums <- function(net) {
  df <- net |> 
      filter(Visits != 0) |> 
      group_by(iter) |> 
      summarise(plants = n_distinct(Plant),
                frugs = n_distinct(Animal),
                net_id = first(net_id))
}

mynum <- lapply(nets, nums)

numbers <- bind_rows(mynum)
```


Read dataset:

```{r}
ind.spec <- readRDS(here("data/WIC_TNW_posts.rds")) |> 
  left_join(numbers)
```

```{r}
summarytools::dfSummary(ind.spec)

summarytools::descr(ind.spec$WIC.TNW)

ind.spec |> 
  summarise(bayestestR::ci(WIC.TNW, ci = 0.90))
```


Load colors:

```{r include=FALSE, message=FALSE}
net_cols <- read_csv(here("data/net_colors.csv"))

cols.bio <- as.character(net_cols$cols_bioregions2)
names(cols.bio) <- as.character(net_cols$plant_sp)

cols.bio.d <- as.character(net_cols$cols_bioregions_d2)
names(cols.bio.d) <- as.character(net_cols$bioregion)
```

Load additional info:

```{r include=FALSE, message=FALSE}
ind_nets_ID <- read.csv(here("data/net.level.selection.csv")) |> 
  filter(type == "ind") |> 
  select(net_id, net_n, net_code, code_ID, ref, 
         family, plant_sp, bioregion, continent,
         plot_rank1, plot_rank2, 
         number.of.species.HL, number.of.species.LL)
```

Create rank for later ploting:

```{r}
rank <- ind.spec |> 
  left_join(ind_nets_ID)|> 
  # extract order to plot (1st continent, then WIC.TNW)
  group_by(continent, bioregion, net_id) |> 
  summarise(mean.WT = mean(WIC.TNW)) |> 
  arrange(continent, mean.WT) |> 
  ungroup() |> 
  mutate(rank_cont = 1:n()) |> 
  arrange(desc(bioregion), mean.WT) |> 
  mutate(rank_bio = 1:n())
```

Add network info to dataset:

```{r}
ind.spec2 <- ind.spec |> 
  left_join(rank) |> 
  left_join(ind_nets_ID)
```


## MODEL

```{r}
ind.spec.sum <- ind.spec2 |> 
  group_by(net_id, bioregion) |> 
  summarise(WIC = mean(WIC),
            TNW = mean(TNW),
            plants = mean(plants),
            frugs = mean(frugs)) 

library(glmmTMB)
library(DHARMa)

hist(ind.spec2$WIC)
hist(ind.spec2$TNW)
```

1. Correlation TNW and number of frugivores

```{r}
cor.test(ind.spec.sum$TNW, ind.spec.sum$frugs)
```

2. Relation WIC ~ TNW (log-log)

```{r}
m2 <- glmmTMB(log(WIC) ~ log(TNW) +
                (1|net_id) + (1|ref), 
              family = gaussian,
              data = ind.spec2)

# m2 <- glmmTMB(log(WIC) ~ log(TNW), 
#               family = gaussian,
#               data = ind.spec.sum)

summary(m2)
simulateResiduals(m2, plot = TRUE)
```

3. Differences in WIC/TNW between bioregions

```{r}
m3 <- glmmTMB(WIC.TNW ~  
                bioregion +
                (1|net_id) + (1|ref), 
              family = gaussian,
              data = ind.spec2)

summary(m3)
simulateResiduals(m3, plot = TRUE)

library(modelbased)

estimate_means(m3, contrast = "bioregion", ci = .90)
```

4. Differences in TNW between bioregions

```{r}
m4 <- glmmTMB(TNW ~  
                bioregion +
                (1|net_id), 
              family = gaussian,
              data = ind.spec2)

summary(m4)
simulateResiduals(m4, plot = TRUE)
```


## PLOTS

Plot WIC/TNW:

```{r message=FALSE}
ind.spec.sum <- ind.spec2 |> 
  group_by(net_id, net_code, net_n, continent, bioregion, plant_sp) |> 
  summarise(rank_cont = first(rank_cont),
            rank_bio = first(rank_bio),
            WIC.TNW = mean(WIC.TNW)) 
  
ggplot(ind.spec2, aes(y = fct_reorder(as.character(net_n), rank_bio, .fun = mean), 
                      x = WIC.TNW, 
                      color = plant_sp, 
                      fill = plant_sp, 
                      label = net_n)) +
  stat_pointinterval(.width = c(0.6, 0.9), point_interval = "mean_qi", 
                     point_size = 3,
                     pch = 21, point_color = "grey20",
                     stroke = 0.5) + 
  geom_text(data = ind.spec.sum,
            aes(y = fct_reorder(as.character(net_n), rank_bio),
                x = WIC.TNW, color = plant_sp,
                label = net_n),
            inherit.aes = FALSE,
            nudge_x = -0.02, 
            nudge_y = 0.6,
            size = 2.5) +
  scale_color_manual(values = cols.bio) + 
  scale_fill_manual(values = cols.bio) + 
  theme(legend.position = "none",
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(),
        panel.background = element_rect(colour = "black"),
        axis.text.x = element_text(size = 12)) + 
  labs(x = NULL, y = NULL) +
  scale_x_continuous(limits = c(0,1)) +
  scale_y_discrete(expand = c(0.05,0))

ggsave(here("figs/Fig_3_WIC_TNW.pdf"), width = 8, height = 5)
```


```{r}
ind.spec2 |> 
  group_by(net_id, continent, bioregion, plant_sp, net_n) |> 
  summarise(WIC = mean(WIC), TNW = mean(TNW), n_frug = first(number.of.species.HL)) |> 
  ggplot(aes(x = TNW, y = WIC, 
             fill = bioregion, 
             size = n_frug)) + 
  geom_point(shape = 21, color = "grey20") +
  ggrepel::geom_text_repel(aes(label = net_n), size = 2) + 
  scale_fill_manual(values = cols.bio.d) + 
  theme(#legend.position = "none",
        panel.border = element_rect(linewidth = 0.5, fill = NA)) + 
  geom_abline(slope = 1, linetype = 3) +
  scale_y_log10() + scale_x_log10() +
  labs(y = "WIC", x  = "TNW", size = "No. frugivores", color = "Region", fill = "Region") 

ggsave(here("figs/Fig_S5_WIC_vs_TNW_log.pdf"), width = 7, height = 5)
```


```{r}
sum <- ind.spec2 |> 
  group_by(net_id, continent, bioregion, plant_sp, net_n) |> 
  summarise(WIC = mean(WIC), TNW = mean(TNW), n_frug = mean(frugs)) 
  
ggplot(sum, aes(x = TNW, y = WIC, 
             fill = bioregion, 
             size = n_frug)) + 
  geom_point(data = ind.spec2, inherit.aes = FALSE,
             aes(x = TNW, y = WIC, color = bioregion), alpha = .02) +
  geom_point(shape = 21, color = "grey20") +
  ggrepel::geom_text_repel(aes(label = net_n), size = 2) + 
  scale_fill_manual(values = cols.bio.d) + 
  scale_color_manual(values = cols.bio.d) + 
  theme(#legend.position = "none",
        panel.border = element_rect(linewidth = 0.5, fill = NA)) + 
  geom_abline(slope = 1, linetype = 3) +
  scale_y_log10() + scale_x_log10() +
  labs(y = "WIC", x  = "TNW", size = "No. frugivores", color = "Region", fill = "Region") 
```


